Honors Draft

### 2020-02-10 chapters/article-draft.Rmd
updates to article draft

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

@@

@@

@@

# Results

@@

Model 1 uses all 18 covariates with simple Normal(0, 1) priors and no overdispersion parameters. The Stan program used to fit Model 1, as well as the other four models, can be found in the appendix. Model 1 fits quickly and efficiently. The Stan output for $\beta_0$ and $\beta$ is below.

\begin{figure}[H]
  \centering
  \includegraphics[width = 5in]{ppcheck_mod1.png}
  \caption{Posterior predictive check for Model 1.}
\end{figure}

Model 1 clearly underestimates the number of block groups with low ridership. This is due in large part to the extra-Poisson variance in Metro Transit's ridership data.

Model 2 builds upon Model 1 by incorporating and overdispersion parameter. Model 2 is *significantly* more computationally expensive to fit than Model 1. This is due to relatively large autocorrelation between each iteration, visible in the traceplots below (Figure ___) as well as in the significantly lower `n_eff` for Model 2. The parameter estimates for $\beta_0$ and $\beta$ for Model 2 are below.

Because we gave $\theta$ and $\sigma$ vague priors, we can see a *much* better fit in the posterior predictive checks for Model 2. One *could* stop analysis here. However, the addition of $\theta \sigma$ only improves the *fit* of the model, not our understanding of the data. Therefore, for this study, we will decompose $\theta \sigma$ into spatial and non-spatial random effects.

First, however, we implement the horseshoe priors to determine if any of our 18 covariates $x_k$ should be removed from the model.

<!-- Notably, Model 3 experienced one divergent transition after warmup. However, the model was fitted several times with no divergent transitions. A "false alarm" divergent transition can occur when the cutoff for determining a divergent transition is too low; this can be true in the case of particularly complex posteriors. Because only 1 out of 20,000 iterations produced a divergent transition, we can proceed with caution using Model 3.  -->

@@

# Appendix

## Model 1
```{r}
writeLines(readLines("~/Documents/honors/honors-project/stan/final-stan/poisson.stan"))
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

@@

### Model 1

We will call this baseline Poisson regression Model 1. Model 1 uses all 18 covariates with simple Normal(0, 1) priors and no overdispersion parameters. The Stan program used to fit Model 1, as well as the other four models, can be found in the appendix. Model 1 fits quickly and efficiently. The Stan output for $\beta_0$ and $\beta$ is below.

\begin{figure}[H]
  \centering
  \includegraphics[width = 6in]{ppcheck_mod1.png}
  \caption{Posterior predictive check for Model 1.}
\end{figure}

Model 1 clearly underestimates the number of block groups with low ridership. This is due in large part to extra-Poisson variance present in Metro Transit's ridership data.

@@

### Model 2

Model 2 builds upon Model 1 by incorporating the set of random effects described above. Model 2 is *significantly* more computationally expensive to fit than Model 1. This is due to relatively large autocorrelation between each iteration, visible in the traceplots below (Figure ___). This autocorrelation made it necessary to run Model 2 on four chains for 20,000 iterations each. Despite this, `n_eff` is significantly lower for Model 2 than Model 1. Reparameterization could likely reduce autocorrelation between iterations.

The parameter estimates for $\beta_0$ and $\beta$ for Model 2 are below.

Because we gave $\theta$ and $\sigma$ vague priors, we can see a *much* better fit in the posterior predictive checks for Model 2. One *could* stop analysis here. However, the addition of $\theta \sigma$ only improves the *fit* of the model, not our understanding of the data. Therefore, for this study, we will decompose $\theta \sigma$ into spatial and non-spatial random effects. First, however, we will implement a Bayesian variable selection process to determine which, if any, of our covariates $x_k$ should be removed from the analysis.

@@

### Model 3

For this study, we implement the regularize horseshoe using code modified from Piironen & Vehtari.

Notably, Model 3 experienced one divergent transition after warmup. However, the model was fitted several times with no divergent transitions. A "false alarm" divergent transition can occur when the cutoff for determining a divergent transition is too low; this can be true in the case of particularly complex posteriors. Because only 1 out of 20,000 iterations produced a divergent transition, we can proceed with caution using Model 3.

@@

### Model 4

Model 4 uses all 18 covariates of interest, no horseshoe priors, and decomposes $\theta\sigma$ into a set of random effects plus a spatial ICAR component. This model converges much more efficiently than Model 2 or Model 3 due to careful reparameterization of the BYM by Riebler et al (2016). Future would could potentially improve the speed of fitting Models 2 and 3.

Model 4 was run on four chains for 10,000 iterations each. The parameter estimates for Model 4 are below. Note that `n_eff` is lower than for Model 1 but significantly higher than Model 2 or 3.

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \toprule
Parameter & Rhat & n\_eff & mean & sd & 2.5\% & 97.5\% \\ 
  \midrule
beta0 & 1.001 & 4770 & -1.880 & 0.028 & -1.937 & -1.825 \\ 
  sigma & 1.003 &  520 & 2.698 & 0.372 & 1.993 & 3.451 \\ 
  rho & 1.002 &  514 & 0.888 & 0.038 & 0.793 & 0.941 \\ 
  theta[5] & 1.000 & 17797 & -0.179 & 0.777 & -1.688 & 1.336 \\ 
  beta[1] & 1.000 & 4322 & 0.034 & 0.047 & -0.056 & 0.127 \\ 
  beta[2] & 1.001 & 3801 & 0.061 & 0.035 & -0.009 & 0.131 \\ 
  beta[3] & 1.000 & 4203 & -0.003 & 0.039 & -0.080 & 0.073 \\ 
  beta[4] & 1.001 & 4330 & -0.049 & 0.044 & -0.134 & 0.037 \\ 
  beta[5] & 1.000 & 4706 & 0.209 & 0.042 & 0.127 & 0.292 \\ 
  beta[6] & 1.000 & 4919 & 0.065 & 0.038 & -0.009 & 0.138 \\ 
  beta[7] & 1.002 & 3929 & 0.029 & 0.038 & -0.046 & 0.104 \\ 
  beta[8] & 1.000 & 6278 & 0.145 & 0.032 & 0.083 & 0.207 \\ 
  beta[9] & 1.000 & 4607 & -0.088 & 0.031 & -0.150 & -0.026 \\ 
  beta[10] & 1.000 & 6145 & 0.014 & 0.029 & -0.043 & 0.070 \\ 
  beta[11] & 1.000 & 6978 & -0.022 & 0.031 & -0.083 & 0.040 \\ 
  beta[12] & 1.000 & 6125 & -0.068 & 0.034 & -0.136 & -0.002 \\ 
  beta[13] & 1.000 & 5875 & 0.094 & 0.033 & 0.029 & 0.158 \\ 
  beta[14] & 1.000 & 4743 & -0.089 & 0.035 & -0.157 & -0.019 \\ 
  beta[15] & 1.001 & 2720 & 0.076 & 0.039 & 0.001 & 0.151 \\ 
  beta[16] & 1.000 & 6053 & -0.029 & 0.031 & -0.091 & 0.031 \\ 
   \bottomrule
\end{tabular}
\caption{Stan output for Model 4} 
\end{table}

Like Models 2 and 3, the BYM2 model performs well in the posterior predictive check (Figure __).

\begin{figure}[H]
  \centering
  \includegraphics[width = 6in]{ppcheckmod4.png}
  \caption{Posterior predictive check for Model 1.}
\end{figure}

### Model 5

Model 5 builds upon Model 4 by incorporating regularized horsehoe priors in addition to spatial ICAR priors. This model has high autocorrelation between iterations and took *significantly* longer than any of the other models to fit in Stan (upwards of 12 hours). That said, the `n_eff` and $\hat{R}$ diagnostics indicate convergence and the parameter estimates (Table __) match our expectations in terms of sign and magnitude based on the previous 4 models.

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \toprule
Parameter & Rhat & n\_eff & mean & sd & 2.5\% & 97.5\% \\ 
  \midrule
beta0 & 1.0 & 5837 & -1.9 & 0.0 & -1.9 & -1.8 \\ 
  sigma & 1.0 & 721 & 2.8 & 0.4 & 2.1 & 3.6 \\ 
  rho & 1.0 & 713 & 0.9 & 0.0 & 0.8 & 0.9 \\ 
  beta[1] & 1.0 & 3637 & -0.1 & 0.1 & -0.2 & -0.0 \\ 
  beta[2] & 1.0 & 8170 & -0.0 & 0.0 & -0.1 & 0.0 \\ 
  beta[3] & 1.0 & 3987 & 0.0 & 0.0 & -0.0 & 0.1 \\ 
  beta[4] & 1.0 & 8531 & -0.0 & 0.0 & -0.1 & 0.0 \\ 
  beta[5] & 1.0 & 7619 & -0.0 & 0.0 & -0.1 & 0.0 \\ 
  beta[6] & 1.0 & 2999 & 0.2 & 0.1 & 0.1 & 0.3 \\ 
  beta[7] & 1.0 & 3726 & 0.0 & 0.0 & -0.0 & 0.1 \\ 
  beta[8] & 1.0 & 7956 & 0.0 & 0.0 & -0.0 & 0.1 \\ 
  beta[9] & 1.0 & 5622 & 0.0 & 0.0 & -0.0 & 0.1 \\ 
  beta[10] & 1.0 & 7530 & 0.1 & 0.0 & 0.1 & 0.2 \\ 
  beta[11] & 1.0 & 4558 & -0.1 & 0.0 & -0.1 & 0.0 \\ 
  beta[12] & 1.0 & 11150 & 0.0 & 0.0 & -0.0 & 0.0 \\ 
  beta[13] & 1.0 & 8289 & -0.0 & 0.0 & -0.1 & 0.0 \\ 
  beta[14] & 1.0 & 5469 & -0.0 & 0.0 & -0.1 & 0.0 \\ 
  beta[15] & 1.0 & 4822 & 0.1 & 0.0 & -0.0 & 0.1 \\ 
  beta[16] & 1.0 & 3863 & -0.1 & 0.0 & -0.1 & 0.0 \\ 
  beta[17] & 1.0 & 2094 & 0.1 & 0.0 & -0.0 & 0.1 \\ 
  beta[18] & 1.0 & 6949 & -0.0 & 0.0 & -0.1 & 0.0 \\ 
   \bottomrule
\end{tabular}
\caption{Stan output for Model 5} 
\end{table}

@@

Naturally, the baseline Model 1 converges most quickly and has the highest `n_eff` diagnostics. The simplicity of the Model, however, yields a poor model fit. Perhaps unexpectedly, Model 4, the BYM2, is the second-most efficient in terms of convergence and Effective Sample Size.

@@

# Appendix: Stan Code

## Model 1

The baseline 
```{r}
writeLines(readLines("~/Documents/honors/honors-project/stan/final-stan/poisson.stan"))
```
